---
title: "Activity Monitoring Report"
author: "Vinod M"
date: "Tuesday, April 14, 2015"
output: html_document
---
##Loading and preprocessing the data :
First the dataset (available as csv) is read into the R environment ```{r, echo=TRUE}data <- read.csv("C:/RLearning/RepResearch/activity.csv")``` . The date column in the dataset is transformed as a factor variable to enable computing the mean, median etc. in the downstream analysis . The totals steps taken per day are summed over the two month period. The histogram of the result is shown below. 


```{r}
data <- read.csv("C:/RLearning/RepResearch/activity.csv")
x <- data
x[,'date'] <- as.factor(x[,'date'])
y <- tapply(x$steps,x$date,sum)
hist(y)
```

The mean of steps taken per day is `r formatC(mean(y,na.rm=TRUE),'%d')` and median is:`r median(y,na.rm=TRUE)`

##Average daily activity pattern
A time series plotof the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis) is shown below:

```{r}
x <- data
x$interval <- as.factor(x$interval)
y <- tapply(x$steps,x$interval,mean,na.rm=TRUE)
 plot(y,type="l",xlab="Interval",ylab="Avg. # of steps",main="Time plot of 5min intervals")
```

The max average steps occur in the `r which.max(y)` interval

##Imputing missing values
Some values in the original dataset are missing the number of steps for some intervals. The total number of missing values are `r table(is.na(data$steps))["TRUE"] ` These values are imputed by using the average value for the interval (calculated above) accross all days
```{r}
y <- data.frame(y)
for(i in 1:nrow(x))
{if(is.na(x[i,1])){x[i,1] <- y[x[i,3],]}}
```

Here is a snapshot of the dataset containing the imputed values:
```{r}
head(x,10)
```

A histogram of the total number of steps taken each day is below:

```{r}
x[,'date'] <- as.factor(x[,'date'])
y <- tapply(x$steps,x$date,sum)
hist(y)
```

The mean of the imputed data is `r mean(y)` and the median is `median(y)`


##Differences in activity patterns between weekdays and weekends?

To infer the differences between weekday and weekend activity, the data is seperated using factors weekday and weekend using a new column (called status) using the R weekdays function applied on the activity date
```{r}

x$date = as.Date(x$date)
y <- weekdays(x$date)
x$status <- ""
for(i in 1:nrow(x))
{if(y[i] =="Saturday" | y[i]=="Sunday"){x[i,"status"] <- "weekend"}else{x[i,"status"] <- "weekday"} }
xx <- list(x$interval,x$status)
y <- tapply(x$steps,xx,sum)
```

A panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis) is shown below :

```{r}

par(mfrow=(c(2,1)))
plot(y[,1],type="l",xlab="Interval",ylab="steps",main="weekday")
plot(y[,2],type="l",xlab="Interval",ylab="steps",main="weekend")
```


